Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-15 3dpas_intro.tex
minor changes, undoing some spelling changes

\author{Shantenu Jha\affil{1}\corrauth\, Daniel S. Katz\affil{2},
  Andre~Luckow\affil{1}, Omer Rana\affil{3}, Yogesh Simmhan\affil{4}, \\ Neil Chue
  Hong\affil{5}}

@@

Analyzing the applications along an analytical space spanned by the three ``D''s
of data-intensive, distributed, and dynamic provides a general and effective
framework.  Subjecting a range of otherwise distinct applications to a common
analytical framework provides advantages to the infrastructure scientists and
cyberinfrastructure experts, namely suggestions for where effort is both needed
and likely to yield maximum returns.  The diverse set of requirements to support
D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.

As D3 applications evolve, our analysis can provide guidance on where to focus
effort on infrastructure.  Last but not least, the advantages of providing a
common but extensible vocabulary that can be used by the entire community,
application end-users, developers and infrastructure scientists alike, in an
emerging field cannot be overstated. One may argue that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with less than desirable consequences.

@@

\item This {\it survey} is not meant to be complete, either in scope of domains
  covered or types of infrastructure and application surveyed.  They by
  definition represent the interests and experience of the authors. The
  requirements and barriers of applications cited in this paper specific are
  mostly representative of the time their representatives were surveyed
  (2012-2013); this is also true of the quantitative analysis of applications.

@@

The (grand) challenge in bio-informatics is the need to provide support
for the analysis of genome sequence data that is becoming available
due to the abundance and high-throughput capability of the
next-generation sequencing devices. The rate of growth of data from
the NGS machines humbles Moore's law; it has increased by more than
three orders of magnitude in barely a decade.

@@

Nowadays, NGS sequencers are even being shipped with servers or small
clusters for performing the alignment and re-sequencing as a pipeline
right after the reads have been generated~\cite{ilm}.
%\alnote{What kind of codes are used
%in these clusters? What are the characteristics of these codes?}
%But any additional analysis is not supported.
%\alnote{What would be an additional analysis}
%
Scientific workflow systems like Taverna~\cite{Taverna:2006}, BioKepler~\cite{biokepler:2011}, and
Galaxy~\cite{galaxy} (though Galaxy does not support distributed systems currently) have been successful in integrating genome
analysis tools for bio-informatics. Script based pipelines that combine
executable are also used often. Very few workflow systems have
explicit support for distributed data or dynamic data handling.
Therefore reusable approaches at extending existing tools to support
distributed and dynamic data are likely to be useful.
%\katznote{Shantenu/Yogesh/Andre, please read this paragraph and tell me if changes
%should be made to this section at this point}\alnote{ok with me}

@@

In addition, scientists can submit Matlab scripts for further custom processing on subsets of
the data that has been integrated. This includes some summarizing processing and plotting
routines.  The only time constraints are on the debugging cycle for debugging the scientist-supplied
reduction codes. These are purely driven by the users' patience.

@@

{\bf MODIS Data Processing (\S\ref{modisKeith})}: Data are brought
into system from various FTP servers.  A pipeline of initial
standardized processing steps on data is done on clouds or HPC
resources.  Scientists can then submit executables and Matlab scripts
that do further custom processing on subsets of the data, which likely
include some summarizing processing (building graphs).  The
application has been run on a cloud and in an HPC environment.

>>>>

\author{Shantenu Jha\affil{1}\corrauth\, Daniel S. Katz\affil{2},
  Andre~Luckow\affil{1}, \\ Omer Rana\affil{3}, Yogesh Simmhan\affil{4}, Neil Chue
  Hong\affil{5}}

@@

Analyzing the applications along an analytical space spanned by the three ``D''s
of data-intensive, distributed, and dynamic provides a general and effective
framework.  Subjecting a range of otherwise distinct applications to a common
analytical framework suggests to infrastructure scientists and
cyberinfrastructure experts where effort is needed
and is likely to yield maximum returns.  The diverse set of requirements to support
D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.

As D3 applications evolve, our analysis can provide guidance on where to focus
effort on infrastructure.  Last but not least, the advantages of providing a
common but extensible vocabulary that can be used by the entire community,
application end-users, developers and infrastructure scientists alike, in an
emerging field cannot be overstated. We believe that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with negative consequences.

@@

\item This {\it survey} is not meant to be complete, either in scope of domains
  covered or types of infrastructure and application surveyed.  They
  represent the interests and experience of the authors. The specific
  requirements and barriers of applications cited in this paper are
  representative of the time at which the applications were surveyed
  (2012-2013); this is also true of the quantitative analysis of applications.

@@

The (grand) challenge in bioinformatics is the need to provide support
for the analysis of genome sequence data that is becoming available
due to the abundance and high-throughput capability of the
next-generation sequencing devices. The rate of growth of data from
the NGS machines humbles Moore's law; it has increased by more than
three orders of magnitude in barely a decade.

@@

Nowadays, NGS sequencers are even being shipped with servers or small
clusters for performing the alignment and re-sequencing as a pipeline
right after the reads have been generated~\cite{ilm}.
%\alnote{What kind of codes are used
%in these clusters? What are the characteristics of these codes?}
%But any additional analysis is not supported.
%\alnote{What would be an additional analysis}
%
Scientific workflow systems like Taverna~\cite{Taverna:2006}, BioKepler~\cite{biokepler:2011}, and
Galaxy~\cite{galaxy} (though Galaxy does not support distributed systems currently) have been successful in integrating genome
analysis tools for bioinformatics. Script-based pipelines that combine
executables are also often used. Very few workflow systems have
explicit support for distributed data or dynamic data handling.
Therefore reusable approaches at extending existing tools to support
distributed and dynamic data are likely to be useful.
%\katznote{Shantenu/Yogesh/Andre, please read this paragraph and tell me if changes
%should be made to this section at this point}\alnote{ok with me}

@@

In addition, scientists can submit Matlab scripts for further custom processing on subsets of
the data that has been integrated. This includes some summarization processing and plotting
routines.  The only time constraints are on the debugging cycle for debugging the scientist-supplied
reduction codes. These are purely driven by the users' patience.

@@

{\bf MODIS Data Processing (\S\ref{modisKeith})}: Data are brought
into system from various FTP servers.  A pipeline of initial
standardized processing steps on data is done on clouds or HPC
resources.  Scientists can then submit executables and Matlab scripts
that do further custom processing on subsets of the data, which likely
include some summarization processing (building graphs).  The
application has been run on a cloud and in an HPC environment.
